---
title: 'Statistical Inference course project  part 2: Basic Inferential Data Analysis'
author: "Mahmoud Nabegh"
date: "December 15, 2018"
output: pdf_document
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
library(ggplot2)
library(datasets)
library(dplyr)
```
## Overview
We use the ToothGrowth data to look at the tooth growth for 10 subjects when taking Orange Juice and Vitamin C with different doses.

## Load the ToothGrowth data

```{r}
library(datasets)
```

## Exploratory analysis

### Data structure
First step is to take a quick look at the data and its structure
```{r}
dim(ToothGrowth)
head(ToothGrowth)
str(ToothGrowth)
```

### Data properties
Next we look at some of the data properties such as:
1. Missing data
```{r}
dim(ToothGrowth)[1]
sum(complete.cases(ToothGrowth))
```
The data contains no missing values

2. Boxplots to represent the 5 number summary for each group by dose

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x = dose, y = len, fill = dose)) + geom_boxplot()
```

So there is very obvious increase in tooth length with increasing the dose.

```{r}
ggplot(ToothGrowth, aes(x = supp, y = len, fill = supp)) + geom_boxplot()
```

Orange juice treatment appears to give a more predictable and better result by having a higher median growth in length and a smaller range, but it is interesting to look at the extreme results of the Vitamin C in case the treatment works better for these cases.

## Hypothesis testing
Here we test the effect of the dose and the method of intaking Vitamin C on the teeth growth

### Dose effect
Our null hypothesis will be that different doses, do not affect the growth length, the data is not paired as they are for different subjects.
```{r}
t.test(ToothGrowth$len[which(ToothGrowth$dose == 1)], ToothGrowth$len[which(ToothGrowth$dose == .5)], paired = F)
```

The test shows the following:

* The p-value < 0.05 which indicates signifacnt result and we reject the null hypothesis.
* The confidence interval does not contain 0 therefore we can infer that a dose of 1 mg is better than 0.5 mg.

```{r}
t.test(ToothGrowth$len[which(ToothGrowth$dose == 2)], ToothGrowth$len[which(ToothGrowth$dose == 1)], paired = F)
```

The results of the 2 mg dose vs the 1 mg dose shows similar results that the 2 mg is better than the 1 mg.

### The method of intake

```{r}
t.test(ToothGrowth$len[which(ToothGrowth$supp == "OJ")], ToothGrowth$len[which(ToothGrowth$supp == "VC")], paired = F)
```

The test shows the following:

* The p-value > 0.05 which indicates non-signifacnt result and we fail to reject the null hypothesis, though it is relatively close to our $\alpha = 0.5$.
* The confidence interval does contain 0 therefore the results is insignificant. 